---
title: "DATA621 Final Project, Does Tanking Work in the NBA?"
author: "Michael O'Donnell"
date: "December 8, 2020"
output: html_document
---

## Load data and required libraries

```{r load_packages echo=FALSE}
# load required packages
library(ggplot2)
library(dplyr)
library(corrplot)
library(MASS)
library(caret)
library(haven)
library(QuantPsyc)
```

```{r import}
# Loading the data
git_dir <- 'https://raw.githubusercontent.com/odonnell31/NBA-Team-Strategies/main/data'
df = read.csv(paste(git_dir, "/nba_teams_data_1990_2020.csv", sep=""))
```

## Exploratory data analysis

See a summary of each variable
```{r}
summary(df)
```

Look at histograms of important predictors
```{r}
# setup 4 plots
par(mfrow=c(2,2))

# plot a histogram for each of the predictor variables
hist(df$consecutive_losing_seasons, main = "Consecutive Losing Seasons")
hist(df$consecutive_playoff_seasons, main = "Consecutive Playoff Seasons")
hist(df$years_since_losing_season, main = "Years Since Last Losing Season")
hist(df$length_of_last_tank, main = "Length of Last Tank (in Seasons)")

```

Subset the data for only possible predictors and response
```{r}
keep_vars <- c("Year", "Team", "Playoffs",
            "Losing_season", "Finals_Team",
            "consecutive_losing_seasons", "consecutive_playoff_seasons",
            "years_since_losing_season", "length_of_last_tank")
df <- df[keep_vars]
```

Check for NA's
```{r}
has_NA = names(which(sapply(df, anyNA)))
has_NA
```

Explore correlations among important predictors
```{r}
# look at correlations 
corr_vars = c("Finals_Team","consecutive_playoff_seasons",
            "years_since_losing_season", "length_of_last_tank")
cor_train = cor(df[corr_vars],  use = "na.or.complete")
corrplot(cor_train)
```

## Creating Regression Model

Create a binary logicstic regression model with Finals_Team as the response
```{r}
# create Binary Logistic Regression model
finals_logistic_model <- glm(Finals_Team ~ consecutive_playoff_seasons +
                              length_of_last_tank +
                              years_since_losing_season,
                              data = df, family = binomial())

summary(finals_logistic_model)
```

Calculate Odd Ratios based on Regression Coefficients
```{r}
#Logistic Regression Coefficient
finals_summary.coeff0 = summary(finals_logistic_model)$coefficient

#Calculating Odd Ratios
FinalsOddRatio = exp(coef(finals_logistic_model))
finals_summary.coeff = cbind(Variable = row.names(finals_summary.coeff0), FinalsOddRatio, finals_summary.coeff0)
row.names(finals_summary.coeff0) = NULL
```

Create a function to standardize the regression coefficients
```{r}
# function to standardize regression coefficients
standardize_coefficients <- function (bl_model) 
{ b <- summary(bl_model)$coef[-1,1]
  sx <- sapply(bl_model$model[-1], sd)
  beta <-(3^(1/2))/pi * sx * b
  return(beta)
}
```

Create standardized regression coefficients with new function
```{r}
# use above function to standardize regression coefficients from model
std_Coeff = data.frame(Standardized.Coeff = standardize_coefficients(finals_logistic_model))
std_Coeff = cbind(Variable = row.names(std_Coeff), std_Coeff)
row.names(std_Coeff) = NULL
```

Merge the Odds Ratios and Coefficients to see all results
```{r}
#Final Summary Report
final_report = merge(finals_summary.coeff, std_Coeff, by = "Variable", all.x = TRUE)

final_report
```